Fluent statistical computing interfaces for biological data analysis

### 2020-04-16 Rmd/01-intro.Rmd
update intro

This thesis focuses on three core aspects of EDA as part of a biological data
science workflow: wrangling, integration and visualisation (in particular we 
focus on the exploration of genomics and transcriptomics data).
It is organised in the same manner that you would iterate back and
forth between these tasks  during an analysis. To begin you wrangle your data 
into a structure that you can understand and use; section \@ref(sec:tidygranges)
introudces a grammar-based framework for transforming genomics data that is 
described in chapter \@ref(ch:plyranges). Next, you integrate data and model 
outputs over genomic regions to gain biological insight; \@ref(sec:ir) 
introduces a framework for incorporating and discovering 'interesting' genomic
regions, described further in chapters \@ref(ch:fluentGenomics) and 
\@ref(ch:intron). Third, you cycle between model-based representations and 
visualisations of your data; section {#sec:va} introduces a visual analytics 
approach for understanding non-linear dimension reduction techniques and is described in
detail in chapter \@ref(ch:tsne). Lastly, you take a step back and look at
the implications and limitations of your work; in chapter \@ref(ch:conclusion)
I describe the outpus of my thesis and plans for future research.

@@

- talking about chapter 5
- taking a step back in the last chapter, have been focused on data wrangling
but am now looking the integration of vis and models
- focussed on a common part of EDA as applied to transcriptomics, dimension
reduction workflows
- especially non-linear dr as that has gotten extremely popular
- combining with the tour, we can grasp the trade-off between global and
local views of the data
- user interaction provides ability to dissect the quality of a DR, and
perform cluster oriented tasks

>>>>

This thesis focuses on core aspects of EDA as part of a biological data
science workflow: wrangling, integration and visualisation, with a focus 
on applications to genomics and transcriptomics. It is organised in 
the same manner that you would iterate back and
forth between these tasks during an analysis. To begin you wrangle your data 
into a structure that you can understand and use; section \@ref(sec:tidygranges)
introduces a grammar-based framework for transforming genomics data that is 
described in chapter \@ref(ch:plyranges). Next, you integrate data and model 
outputs over genomic regions to gain biological insight; \@ref(sec:ir) 
introduces a framework for incorporating and discovering 'interesting' genomic
regions, described further in chapters \@ref(ch:fluentGenomics) and 
\@ref(ch:intron). Third, you cycle between model-based representations and 
visualisations of your data; section \@ref(sec:va) introduces a visual analytics 
approach for understanding non-linear dimension reduction techniques and is 
described in detail in chapter \@ref(ch:tsne). Lastly, you take a step back and 
look at the implications and limitations of your work; in chapter 
\@ref(ch:conclusion) I describe the outputs of my thesis and plans for future 
research.

@@

In the final chapter, we move away from data wrangling and towards the 
integration of visualisation with model-based summaries of high-dimensional
data sets. We focus on a common tool for EDA (especially applied to single-cell
transcriptomics): dimension reduction (DR). We consider the incorporation of 
interactive and dynamic graphics to assist analysts in using DR techniques
for cluster orientation tasks. In particular, we advocate for the use of tours
[@Cook1995-bi] alongside a DR visualisation to highlight potential pitfalls and 
distortions of obtained from a DR method. This approach acknowledges that
there is no 'one' best visualisation or DR for a high-dimensional dataset,
and we often want to have an understanding of both the global and local structure
within our data.

Chapter \@ref(ch:liminal) introduces
a software package called *liminal* for constructing these views and
a user interaction framework for identifying distortions. We present several
case studies for using our approach to diagnose the quality of 
results obtained via t-distributed stochastic neighbour embeddings (t-SNE)
